---
title: "ISSPxVTS_v5 analysis"
author: "Yu-Chin Chiu"
date: "July 11, 2018"
output: github_document
---


```{r knitSetup}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='figure/', echo = FALSE, warning = FALSE, message = FALSE)
options(width = 9999)

```

#### First, Set up the environment and load the data: gpData.csv 
This file came from python scripts - cleaning and preprocessing

```{r setup}
rm(list=ls())
library('tidyverse')
library('ez')
library('apaTables')
library('BayesFactor')
currentDir <- getwd()
gpData = read_csv('gpData.csv',col_names=TRUE)
source("getWSSE.R")

```


#### Set up some formating for the plots
Set up some good format, here I am using apatheme with white background, with black axis lines, no grids.

```{r fig apa-format}
apatheme = theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line=element_line())
dodge = position_dodge(width=0.9)


```


```{r analysis_prep}
# create a new variable called "run_3bk": run3,4 = B1, run5-6 = B2, run7-8 = B3

gpData[which(gpData$phase=="hybrid" & gpData$runId==3), "run_3bk"]="B1"
gpData[which(gpData$phase=="hybrid" & gpData$runId==4), "run_3bk"]="B1"
gpData[which(gpData$phase=="hybrid" & gpData$runId==5), "run_3bk"]="B2"
gpData[which(gpData$phase=="hybrid" & gpData$runId==6), "run_3bk"]="B2"
gpData[which(gpData$phase=="hybrid" & gpData$runId>=7), "run_3bk"]="B3"

gpData[which(gpData$phase=="hybrid" & gpData$runId<=5), "run_2bk"]="first3"
gpData[which(gpData$phase=="hybrid" & gpData$runId>=6), "run_2bk"]="last3"


gpData[which(gpData$phase=="hybrid" & gpData$runId<=4), "runId2"]="first2"
gpData[which(gpData$phase=="hybrid" & gpData$runId>=5), "runId2"]="last4"


# mark trials with ACC=0, RT=NA, only analyze CORRECT TRIALS' RT
gpData[which(gpData$sbjACC==0), "sbjRT"] =NA
gpData[which(gpData$sbjRT<250), "sbjRT"] =NA
gpData[which(gpData$sbjRT>1500), "sbjRT"]=NA

# Turn a bunch of codings into categorical factors
catVblList <- c("phase","sbjId","runId", "swProb", "task", "trialType_sw", "run_3bk","run_2bk","runId2")
idx <- match(catVblList, colnames(gpData))
for (c in idx[!is.na(idx)]){
  gpData[[c]] <- factor(gpData[[c]], ordered = TRUE)
}

gpData$phase <- factor(gpData$phase, levels = c("training", "hybrid"), ordered=TRUE)
gpData$trialType <- factor(gpData$trialType, levels = c("cued", "choice"), ordered=TRUE)
gpData$trialType_sw <- factor(gpData$trialType_sw, levels = c("repeat", "switch"), ordered=TRUE)


NrS <- length(unique(gpData$sbjId))
NrS

```

```{r}
# exclude subjects with low task-switching accuracy and very low VSR
mean_tsACC_cuedTrials <- gpData %>% 
  filter(trialType=="cued") %>%
  group_by(sbjId) %>%
  summarise(tsACC = mean(sbjACC, na.rm = TRUE))

mean_tsACC_training <- gpData %>% 
  filter(phase=="training") %>%
  group_by(sbjId) %>%
  summarise(tsACC = mean(sbjACC, na.rm = TRUE))

mean_VSR <- gpData %>% 
  filter(phase=="hybrid" & trialType=="choice") %>%
  group_by(sbjId) %>%
  summarise(VSR = mean(trialType_int, na.rm = TRUE))  

valid_sbjList <- c(NULL)
cnt=0
for (i in 1:nrow(mean_VSR)){
  if (mean_tsACC_training$tsACC[i] < .70 | mean_VSR$VSR[i] < .05) {
  }
  else{
    cnt=cnt+1
    valid_sbjList[cnt]<- mean_tsACC_training$sbjId[i]
  }
}

NrS_valid <- length(valid_sbjList)

NrS_valid

gpData_raw <-gpData
gpData<- tibble()
gpData <- gpData_raw[gpData_raw$sbjId %in% valid_sbjList, ] 

```
## RESULTS: 
### 1. Cued trials 
#### Analyze RT & ACC with 2 phase x 2 swProb x trialType_sw 
```{r}
# create subject level condition means
CondM1 <- tibble()
CondM1 <- gpData %>% 
  filter(trialType=="cued") %>%
  group_by(sbjId, phase, swProb, trialType_sw) %>%
  summarise(meanRT = mean(sbjRT, na.rm = TRUE), meanACC = mean(sbjACC, na.rm = TRUE)) %>%  ungroup() 


# 3-way ANOVA: cued trials: RT ~ 2 phase x 2 swProb x 2 trialType_sw
z = ezANOVA(data = CondM1
            , dv = meanRT
            , wid = sbjId
            , within = .(phase, swProb, trialType_sw)
            , within_full = .(phase, swProb, trialType_sw)
            , type = 3
            , detailed = TRUE)
z_table <- apa.ezANOVA.table(z, filename="RT_2phasex2swProbx2trialTypeSw.doc")
print(z_table)


# 3-way ANOVA: cued trials: ACC ~ 2 phase x 2 swProb x 2 trialType_sw
z = ezANOVA(data = CondM1
            , dv = meanACC
            , wid = sbjId
            , within = .(phase, swProb, trialType_sw)
            , within_full = .(phase, swProb, trialType_sw)
            , type = 3
            , detailed = TRUE)
z_table <- apa.ezANOVA.table(z, filename="ACC_2phasex2swProbx2trialTypeSw.doc")
print(z_table)


# prepare for the figure, calculate within-S SEM
df <- CondM1 %>% select(-meanACC)
colnames(df)[colnames(df)=="meanRT"] <- "M"
LnMSE_rt <- getWSSE(df)

df <- CondM1 %>% select(-meanRT)
colnames(df)[colnames(df)=="meanACC"] <- "M"
LnMSE_acc <- getWSSE(df)


# create group level condition means
gpCondM1 <- CondM1 %>%
  group_by(phase, swProb, trialType_sw) %>% 
  summarise(gpmeanRT = mean(meanRT), SE_rt = LnMSE_rt , gpmeanACC = mean(meanACC)*100, SE_acc = LnMSE_acc*100) 
print(gpCondM1)


```

#### Figure 1a: Cued - RT
```{r Figure1a}
gpCondM1 %>%
  ggplot(aes(x=swProb, y = gpmeanRT, fill = trialType_sw)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanRT + (1*SE_rt), ymin =gpmeanRT -(1*SE_rt)), position = dodge, width=.25) + facet_grid(.~phase) + 
  ylab("mean RT (correct trials)") + 
  coord_cartesian(ylim = c(600,900)) +
  scale_fill_grey()
```
#### Figure 1b: Cued - ACC
```{r Figure1b}
gpCondM1 %>%
  ggplot(aes(x=swProb, y = gpmeanACC, fill = trialType_sw)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanACC + (1*SE_acc), ymin =gpmeanACC -(1*SE_acc)), position = dodge, width=.25) + facet_grid(.~phase) + 
  ylab("mean ACC") + 
  coord_cartesian(ylim = c(80,100)) +
  scale_fill_grey()
```


### 2. Choice trials
#### Analyze RT & ACC with 2 swProb x 2 trialType_sw 
```{r}

# create subject level condition means
CondM2 <- tibble()
CondM2 <- gpData %>% 
  filter(trialType=="choice") %>%
  filter(trialType_sw=="switch" | trialType_sw=="repeat") %>%
  group_by(sbjId, swProb, trialType_sw) %>%
  summarise(meanRT = mean(sbjRT, na.rm = TRUE), meanACC = mean(sbjACC, na.rm = TRUE)) %>% ungroup() 


# 2-way ANOVA: choice trials: RT ~ 2 swProb x 2 trialType_sw
z = ezANOVA(data = CondM2
            , dv = meanRT
            , wid = sbjId
            , within = .(swProb, trialType_sw)
            , within_full = .(swProb, trialType_sw)
            , type = 3
            , detailed = TRUE)
z_table <- apa.ezANOVA.table(z, filename="Choice_RT_2swProbx2trialTypeSw.doc")
print(z_table)


# 2-way ANOVA: choice trials: ACC~ 2 swProb x 2 trialType_sw
z = ezANOVA(data = CondM2
            , dv = meanACC
            , wid = sbjId
            , within = .(swProb, trialType_sw)
            , within_full = .(swProb, trialType_sw)
            , type = 3
            , detailed = TRUE)
z_table <- apa.ezANOVA.table(z, filename="Choice_ACC_2swProbx2trialTypeSw.doc")
print(z_table)


# prepare for the figure, calculate within-S SEM
df <- CondM2 %>% select(-meanACC)
colnames(df)[colnames(df)=="meanRT"] <- "M"
LnMSE_rt <- getWSSE(df)

df <- CondM2 %>% select(-meanRT)
colnames(df)[colnames(df)=="meanACC"] <- "M"
LnMSE_acc <- getWSSE(df)


# create group level condition means
gpCondM2 <- CondM2 %>%
  group_by(swProb, trialType_sw) %>% 
  summarise(gpmeanRT = mean(meanRT), SE_rt = LnMSE_rt , gpmeanACC = mean(meanACC)*100, SE_acc = LnMSE_acc*100) 
print(gpCondM2)


```


#### Figure 2a: Choice - RT
```{r Figure2a, fig.height = 3, fig.width = 5}
gpCondM2 %>%
  ggplot(aes(x=swProb, y = gpmeanRT, fill = trialType_sw)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanRT + (1*SE_rt), ymin =gpmeanRT -(1*SE_rt)), position = dodge, width=.25)+ 
  ylab("mean RT (correct trials)") + 
  coord_cartesian(ylim = c(600,900)) +
  scale_fill_grey()
```


#### Figure 2b: Choice - ACC
```{r Figure2b, fig.height = 3, fig.width = 5}
gpCondM2 %>%
  ggplot(aes(x=swProb, y = gpmeanACC, fill = trialType_sw)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanACC + (1*SE_acc), ymin =gpmeanACC -(1*SE_acc)), position = dodge, width=.25) + 
  ylab("mean ACC") + 
  coord_cartesian(ylim = c(85,100)) +
  scale_fill_grey()
```


### 3. Choice trials
#### Analyze VSR and tskRatio with 2 swProb 
```{r}

# create subject level condition means
CondM3 <- tibble()
CondM3 <- gpData %>% 
  filter(trialType=="choice") %>%
  filter(trialType_sw=="switch" | trialType_sw=="repeat") %>%
  group_by(sbjId, swProb) %>%
  summarise(meanVSR = mean(trialType_int, na.rm = TRUE), meanTskRatio = mean(task_int, na.rm = TRUE)) %>% ungroup() 


# 1-way ANOVA: choice trials: VSR ~ 2 swProb 
z = ezANOVA(data = CondM3
            , dv = meanVSR
            , wid = sbjId
            , within = .(swProb)
            , within_full = .(swProb)
            , type = 3
            , detailed = TRUE)
z_table <- apa.ezANOVA.table(z, filename="VSR_2swProb.doc")
print(z_table)


# 1-way ANOVA: choice trials: tskRatio ~ 2 swProb 
z = ezANOVA(data = CondM3
            , dv = meanTskRatio
            , wid = sbjId
            , within = .(swProb)
            , within_full = .(swProb)
            , type = 3
            , detailed = TRUE)
z_table <- apa.ezANOVA.table(z, filename="tskRatio_2swProb.doc")
print(z_table)

# prepare for the figure, calculate within-S SEM
df <- CondM3 %>% select(-meanTskRatio)
colnames(df)[colnames(df)=="meanVSR"] <- "M"
LnMSE_VSR <- getWSSE(df)

df <- CondM3 %>% select(-meanVSR)
colnames(df)[colnames(df)=="meanTskRatio"] <- "M"
LnMSE_tskR <- getWSSE(df)


# create group level condition means
gpCondM3 <- CondM3 %>%
  group_by(swProb) %>% 
  summarise(gpmeanVSR = mean(meanVSR)*100, SE_vsr = LnMSE_VSR*100 , gpmeanTskRatio = mean(meanTskRatio)*100, SE_tskR = LnMSE_tskR*100) 
print(gpCondM3)


```


#### Figure 3a: VSR ~ swProb
```{r Figure3a, fig.height = 4, fig.width = 3}

# Use boxplot here.. need to provide subject level tibble
CondM3 %>%
  ggplot(aes(x=swProb, y = meanVSR)) + 
  geom_boxplot(alpha = 0) +
  geom_jitter(alpha = 0.3, color = "tomato") +
  xlab('swProb')+
  ylab("mean VSR") + 
  ggtitle("mean voluntary switch rate") 

```


#### Figure 3b: taskBias ~swProb
```{r Figure3b, fig.height = 4, fig.width = 3}

# Use boxplot here.. need to provide subject level tibble
CondM3 %>%
  ggplot(aes(x=swProb, y = meanTskRatio)) + 
  geom_boxplot(alpha = 0) +
  geom_jitter(alpha = 0.3, color = "tomato") +
  xlab('swProb')+
  ylab("mean % task 1 chosen") + 
  ggtitle("Any task Bias") 
```


# Exploratory analysis


### 4. Choice trials
#### Analyze VSR  with 3 blocks ("run_3bk") x 2 swProb 
```{r}

# create subject level condition means
CondM4 <- tibble()
CondM4 <- gpData %>% 
  filter(trialType=="choice") %>%
  filter(trialType_sw=="switch" | trialType_sw=="repeat") %>%
  group_by(sbjId, run_3bk, swProb) %>%
  summarise(meanVSR = mean(trialType_int, na.rm = TRUE)) %>% ungroup() 


# 2-way ANOVA: choice trials: VSR ~ 3 blocks x 2 swProb 
z = ezANOVA(data = CondM4
            , dv = meanVSR
            , wid = sbjId
            , within = .(run_3bk, swProb)
            , within_full = .(run_3bk, swProb)
            , type = 3
            , detailed = TRUE)
print(z)
z_table <- apa.ezANOVA.table(z, filename="VSR_3bkx2SwProb.doc")


# prepare for the figure, calculate within-S SEM
df <- CondM4
colnames(df)[colnames(df)=="meanVSR"] <- "M"
LnMSE_VSR4 <- getWSSE(df)


# create group level condition means
gpCondM4 <- CondM4 %>%
  group_by(run_3bk, swProb) %>% 
  summarise(gpmeanVSR = mean(meanVSR)*100, SE_vsr = LnMSE_VSR4*100) 
print(gpCondM4)

```


#### Figure 4: VSR: 3 blocks x 2 swProb
```{r Figure4}
gpCondM4 %>%
  ggplot(aes(x=run_3bk, y = gpmeanVSR, fill = swProb)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanVSR + (1*SE_vsr), ymin =gpmeanVSR -(1*SE_vsr)), position = dodge, width=.25) +
  ylab("mean VSR") + 
  coord_cartesian(ylim = c(30,40)) +
  scale_fill_grey()
```


### 5. Choice trials
#### Analyze VSR  with 2 blocks ("run_2bk") x 2 swProb 
```{r}

# create subject level condition means
CondM5 <- tibble()
CondM5 <- gpData %>% 
  filter(trialType=="choice") %>%
  filter(trialType_sw=="switch" | trialType_sw=="repeat") %>%
  group_by(sbjId, run_2bk, swProb) %>%
  summarise(meanVSR = mean(trialType_int, na.rm = TRUE)) %>% ungroup() 


# 2-way ANOVA: choice trials: VSR ~ 2 block x 2 swProb 
z = ezANOVA(data = CondM5
            , dv = meanVSR
            , wid = sbjId
            , within = .(run_2bk, swProb)
            , within_full = .(run_2bk, swProb)
            , type = 3
            , detailed = TRUE)
print(z)
z_table <- apa.ezANOVA.table(z, filename="VSR_2bkx2SwProb.doc")


# prepare for the figure, calculate within-S SEM
df <- CondM5
colnames(df)[colnames(df)=="meanVSR"] <- "M"
LnMSE_VSR5 <- getWSSE(df)


# create group level condition means
gpCondM5 <- CondM5 %>%
  group_by(run_2bk, swProb) %>% 
  summarise(gpmeanVSR = mean(meanVSR)*100, SE_vsr = LnMSE_VSR5*100) 
print(gpCondM5)

```


#### Figure 5: VSR: 2 blocks x 2 swProb
```{r Figure5, fig.width=5, fig.height=3.5}
gpCondM5 %>%
  ggplot(aes(x=run_2bk, y = gpmeanVSR, fill = swProb)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanVSR + (1*SE_vsr), ymin =gpmeanVSR -(1*SE_vsr)), position = dodge, width=.25) +
  ylab("mean VSR") + 
  coord_cartesian(ylim = c(30,40)) +
  scale_fill_grey()
```


### 6. Choice trials
#### Analyze VSR  with 2 blocks ("runId2" - first 2 & last 4) x 2 swProb 
```{r}

# create subject level condition means
CondM6 <- tibble()
CondM6 <- gpData %>% 
  filter(trialType=="choice") %>%
  filter(trialType_sw=="switch" | trialType_sw=="repeat") %>%
  group_by(sbjId, runId2, swProb) %>%
  summarise(meanVSR = mean(trialType_int, na.rm = TRUE)) %>% ungroup() 


# 2-way ANOVA: choice trials: VSR ~ 2 block x 2 swProb 
z = ezANOVA(data = CondM6
            , dv = meanVSR
            , wid = sbjId
            , within = .(runId2, swProb)
            , within_full = .(runId2, swProb)
            , type = 3
            , detailed = TRUE)
print(z)
z_table <- apa.ezANOVA.table(z, filename="VSR_2bk_unevenx2SwProb.doc")


# prepare for the figure, calculate within-S SEM
df <- CondM6
colnames(df)[colnames(df)=="meanVSR"] <- "M"
LnMSE_VSR6 <- getWSSE(df)


# create group level condition means
gpCondM6 <- CondM6 %>%
  group_by(runId2, swProb) %>% 
  summarise(gpmeanVSR = mean(meanVSR)*100, SE_vsr = LnMSE_VSR6*100) 
print(gpCondM6)

```


#### Figure 6: VSR: 2 blocks ("runId2" - first 2 & last 4)  x 2 swProb
```{r Figure6, fig.width=5, fig.height=3.5}
gpCondM6 %>%
  ggplot(aes(x=runId2, y = gpmeanVSR, fill = swProb)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanVSR + (1*SE_vsr), ymin =gpmeanVSR -(1*SE_vsr)), position = dodge, width=.25) +
  ylab("mean VSR") + 
  coord_cartesian(ylim = c(30,40)) +
  scale_fill_grey()
```


### 7. Choice trials
#### Correlation between ISSP_train and overall VSR
```{r}

## calculate ISSP for the training portion each subject has 4 condition means
condM7 <- gpData %>% 
  filter(phase=="training") %>%
  group_by(sbjId, swProb, trialType_sw) %>%
  summarise(meanRT =mean(sbjRT, na.rm = TRUE)) %>%  ungroup() 


# need to spread these factor Columns into numbers to create unique condition id and calculate ISSP_train
nonFactor <- c("sbjId", "meanRT")
colNames <- colnames(condM7)
idx <- match(colNames,nonFactor)
factorCol <- which(is.na(idx))
arg <- ""
c   <- 0
for(f in factorCol){
  condM7[[f]] <- as.character(condM7[[f]])
  c <- c+1
  arg[c] <- paste("condM7$",colNames[f],sep = "")
}
cmd <- paste0("paste0(", paste(arg,collapse = ","),")") 
condM7$name <-eval(parse(text=cmd))

condM7 <- condM7 %>% 
  select(sbjId, meanRT, name)
summary <- spread(condM7,key = name,value = meanRT)

summary$ISSP_train <- (summary$`sw25%switch`-summary$`sw25%repeat`)-(summary$`sw75%switch`-summary$`sw75%repeat`)
summary <- summary %>%
  select(sbjId, ISSP_train)

## calculate mean VSR during hybird
condM7_hybrid <- gpData %>% 
  filter(phase=="hybrid" & trialType=="choice") %>%
  group_by(sbjId) %>%
  summarise(meanVSR = mean(trialType_int, na.rm = TRUE)) %>%  ungroup() 

summary$meanVSR <- condM7_hybrid$meanVSR

```


#### Figure 7: correlation between ISSP_train & mean VSR 
```{r Figure7, fig.height=4, fig.width=4}
summary %>% ggplot(aes(x = ISSP_train, y = meanVSR))+
  geom_point(shape=1) + 
  geom_smooth(method=lm) + 
  ggtitle("Regress meanVSR during hybrid phase against ISSP during training phase")


cor.test(summary$ISSP_train,summary$meanVSR) 

```


### 8. Choice trials
#### Correlation between ISSP_choice and VSReffect
```{r}

## calculate ISSP for the hybrid, choice trials - each subject has 4 condition means
condM8 <- gpData %>% 
  filter(phase=="hybrid" & trialType=="choice") %>%
  filter(trialType_sw=="switch" | trialType_sw=="repeat") %>%
  group_by(sbjId, swProb, trialType_sw) %>%
  summarise(meanRT =mean(sbjRT, na.rm = TRUE)) %>%  ungroup() 


# need to spread these factor Columns into numbers to create unique condition id and calculate ISSP_hybrid_choice
nonFactor <- c("sbjId", "meanRT")
colNames <- colnames(condM8)
idx <- match(colNames,nonFactor)
factorCol <- which(is.na(idx))
arg <- ""
c   <- 0
for(f in factorCol){
  condM8[[f]] <- as.character(condM8[[f]])
  c <- c+1
  arg[c] <- paste("condM8$",colNames[f],sep = "")
}

cmd <- paste0("paste0(", paste(arg,collapse = ","),")") 
condM8$name <-eval(parse(text=cmd))

condM8 <- condM8 %>% 
  select(sbjId, meanRT, name)
tt <- spread(condM8,key = name,value = meanRT)

# use the summary tibble from above 
summary$ISSP_choice <- (tt$`sw25%switch`-tt$`sw25%repeat`)-(tt$`sw75%switch`-tt$`sw75%repeat`)


## calculate mean VSR75 vs. VSR25 during hybird
condM8_tmp <- gpData %>% 
  filter(phase=="hybrid" & trialType=="choice") %>%
  group_by(sbjId, swProb) %>%
  summarise(meanVSR = mean(trialType_int, na.rm = TRUE)) %>%  ungroup() 

condM8_tmp <- spread(condM8_tmp,key = swProb,value = meanVSR)

summary$VSReffect <- condM8_tmp$`sw75%`-condM8_tmp$`sw25%`


```


#### Figure 8: correlation between ISSP_hybrid_choice & VSR effect 
```{r Figure8, fig.height=4, fig.width=4}

summary %>% ggplot(aes(x = ISSP_choice, y = VSReffect))+
  geom_point(shape=1) + 
  geom_smooth(method=lm) + 
  ggtitle("Regress VSR effect (VSR75 vs. VSR25) during hybrid phase against ISSP (choice trials) during hybrid phase")

cor.test(summary$ISSP_choice,summary$VSReffect) 


```


