---
title: "ISSPxVTS_v5 analysis"
author: "Yu-Chin Chiu"
date: "June 11, 2018"
output: md_document
---


```{r knitSetup}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='figure/', echo = FALSE, warning = FALSE, message = FALSE)

```

#### First, Set up the environment and load the data: gpData.csv 
This file came from python scripts - cleaning and preprocessing

```{r setup}
rm(list=ls())
library('tidyverse')
library('ez')

currentDir <- getwd()
gpData = read_csv('gpData.csv',col_names=TRUE)
source("getWSSE.R")

```


#### Set up some formating for the plots
Set up some good format, here I am using apatheme with white background, with black axis lines, no grids.

```{r fig apa-format}
apatheme = theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line=element_line())
dodge = position_dodge(width=0.9)
```


```{r analysis_prep}
# create a new variable called "run_half": run 3-5: FirstHalf, run 6-8: secondHalf

gpData[which(gpData$phase=="hybrid" & gpData$runId<=5), "runId_half"]="firstHalf"
gpData[which(gpData$phase=="hybrid" & gpData$runId>=6), "runId_half"]="SecondHalf"

# mark trials with ACC=0, RT=NA, only analyze CORRECT TRIALS' RT
gpData[which(gpData$sbjACC==0), "sbjRT"]=NA

# Turn a bunch of codings into categorical factors
catVblList <- c("phase","runId","sbjId", "swProb", "task", "trialType_sw", "runId_half")
idx <- match(catVblList, colnames(gpData))
for (c in idx[!is.na(idx)]){
  gpData[[c]] <- factor(gpData[[c]], ordered = TRUE)
}

gpData$phase <- factor(gpData$phase, levels = c("training", "hybrid"), ordered=TRUE)
gpData$trialType <- factor(gpData$trialType, levels = c("cued", "choice"), ordered=TRUE)

NrS <- length(unique(gpData$sbjId))

```

```{r}
# exclude subjects with low task-switching accuracy and very low VSR
mean_tsACC_training <- gpData %>% 
  filter(phase=="training") %>%
  group_by(sbjId) %>%
  summarise(tsACC = mean(sbjACC, na.rm = TRUE))

mean_VSR <- gpData %>% 
  filter(phase=="hybrid" & trialType=="choice") %>%
  group_by(sbjId) %>%
  summarise(VSR = mean(trialType_int, na.rm = TRUE))  

valid_sbjList <- c(NULL)
cnt=0
for (i in 1:nrow(mean_VSR)){
  if (mean_tsACC_training$tsACC[i] < .70 | mean_VSR$VSR[i] < .10) {
  }
  else{
    cnt=cnt+1
    valid_sbjList[cnt]<- mean_tsACC_training$sbjId[i]
  }
}

NrS_valid <- length(valid_sbjList)

```

#### First summarize individual subject's data

```{r}

gpCondM1 <- tibble()
gpCondM2 <- tibble()
gpCondM3 <- tibble()
gpCondM4 <- tibble()
gpCondM5 <- tibble()

for (S in valid_sbjList) {
  mydata <-  gpData %>% filter(sbjId==S) 
  
  
  # 2 phase (training, hybrid) x 2 swProb (25%, 75%) x 2 trialType_sw (switch, repeat) for only the cued trials
  condM <- mydata %>%
    filter(trialType=="cued")  %>%
    group_by(phase, swProb, trialType_sw) %>% 
    summarise(meanRT = mean(sbjRT, na.rm = TRUE), meanACC = mean(sbjACC, na.rm = TRUE))
  condM <- bind_cols(tibble(sbjId = rep(S,dim(condM)[1])), condM)
  ISSP_training <- condM[2,"meanRT"] - condM[1,"meanRT"] - (condM[4,"meanRT"]-condM[3,"meanRT"])
  ISSP_hybrid_cued <- condM[6,"meanRT"] - condM[5,"meanRT"] - (condM[8,"meanRT"]-condM[7,"meanRT"])
  gpCondM1 <- bind_rows(gpCondM1, condM)
  
  
  
  # 2 trialType(cued, choice) x 2 swProb (25%, 75%) x 2 trialType_sw (switch, repeat)
  condM2 <- mydata %>%
    group_by(trialType, swProb, trialType_sw) %>% 
    summarise(meanRT = mean(sbjRT, na.rm = TRUE), meanACC = mean(sbjACC, na.rm = TRUE))
  condM2 <- bind_cols(tibble(sbjId = rep(S,dim(condM2)[1])), condM2)
  gpCondM2 <- bind_rows(gpCondM2, condM2)
  
  
  
  
  # VSR: Choice data: 2 swProb (25%, 75%)
  condM3 <- mydata %>%
  filter(trialType=="choice")  %>%
    group_by(swProb) %>% 
    summarise(meanVSR = mean(trialType_int, na.rm = TRUE), meanACC = mean(sbjACC, na.rm = TRUE), meanRT = mean(sbjRT, na.rm = TRUE),  tskRatio = mean(task_int, na.rm = TRUE))
  condM3 <- bind_cols(tibble(sbjId = rep(S,dim(condM3)[1])), condM3)
  gpCondM3 <- bind_rows(gpCondM3, condM3)
  
  # VSR: Choice data: 2 runHalf(run3-5, run6-8) x 2 swProb (25%, 75%)
  condM4 <- mydata %>%
    filter(trialType=="choice") %>%
    group_by(runId_half, swProb) %>%
    summarise(meanVSR = mean(trialType_int, na.rm = TRUE), meanACC = mean(sbjACC, na.rm = TRUE), meanRT = mean(sbjRT, na.rm = TRUE),  tskRatio = mean(task_int, na.rm = TRUE))
  condM4 <- bind_cols(tibble(sbjId = rep(S,dim(condM4)[1])), condM4)
  gpCondM4 <- bind_rows(gpCondM4, condM4)
  
  # task biases
  condM5 <- mydata %>%
    filter(trialType=="choice") %>%
    group_by(swProb) %>%  
    summarise(tskRatio = mean(task_int))
  condM5 <- bind_cols(tibble(sbjId = rep(S,dim(condM5)[1])), condM5)
  gpCondM5 <- bind_rows(gpCondM5, condM5)


}

```

#### Then, calculate the group mean and within-subject SEM using "getWSSE" (a function I wrote)

```{r ISSP analysis}

# ISSP: cued trials: 2 phases x 2 swProb x 2 trialTpe_sw

df <- gpCondM1 %>% select(-meanACC)
colnames(df)[colnames(df)=="meanRT"] <- "M"
LnMSE_rt <- getWSSE(df)


df <- gpCondM1 %>% select(-meanRT)
colnames(df)[colnames(df)=="meanACC"] <- "M"
LnMSE_acc <- getWSSE(df)


ISSP <- gpCondM1 %>%
  group_by(phase, swProb, trialType_sw) %>% 
  summarise(gpmeanRT = mean(meanRT), SE_rt = LnMSE_rt , gpmeanACC = mean(meanACC)*100, SE_acc = LnMSE_acc*100) 

swCost_75 = ISSP[4,"gpmeanRT"]-ISSP[3,"gpmeanRT"]
swCost_25 = ISSP[2,"gpmeanRT"]-ISSP[1,"gpmeanRT"]

print('Group means:')
print(paste("cued: swCost:75% =", sprintf("%.2f", swCost_75), "ms"))
print(paste("cued: swCost:25% =", sprintf("%.2f", swCost_25), "ms"))

```


### RESULTS: 
#### Figure 1: cued trials: ISSP(phase x swProb x trialtype_sw) effect in **RT** (correct trials)
```{r Figure1}
ISSP %>%
  ggplot(aes(x=swProb, y = gpmeanRT, fill = trialType_sw)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanRT + (1*SE_rt), ymin =gpmeanRT -(1*SE_rt)), position = dodge, width=.25) + facet_grid(.~phase) + 
  ylab("mean RT (correct trials)") + 
  coord_cartesian(ylim = c(600,900)) +
  scale_fill_grey()

```

#### 3-way ANOVA: cued trials: RT ~ 2 phase x 2 swProb x 2 trialType_sw
```{r}
z = ezANOVA(data = gpCondM1
            , dv = meanRT
            , wid = sbjId
            , within = .(phase, swProb, trialType_sw)
            , within_full = .(phase, swProb, trialType_sw)
            , type = 3
            , detailed = TRUE)
print(z)

```

#### 2-way ANOVA: cued trials(training): RT ~ 2 swProb x 2 trialType
```{r}
z = ezANOVA(data = gpCondM1[gpCondM1$phase=="training",]
            , dv = meanRT
            , wid = sbjId
            , within = .(swProb, trialType_sw)
            , within_full = .(swProb, trialType_sw)
            , type = 3
            , detailed = TRUE)
print(z)

```

#### 2-way ANOVA: cued trials(hybrid): RT ~ 2 swProb x 2 trialType
```{r}
z = ezANOVA(data = gpCondM1[gpCondM1$phase=="hybrid",]
            , dv = meanRT
            , wid = sbjId
            , within = .(swProb, trialType_sw)
            , within_full = .(swProb, trialType_sw)
            , type = 3
            , detailed = TRUE)
print(z)

```

#### Figure 2: ISSP(swProb x trialtype_sw) effect in **accuracy** by phase
```{r Figure2}
ISSP %>%
  ggplot(aes(x=swProb, y = gpmeanACC, fill = trialType_sw)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanACC + (1*SE_acc), ymin =gpmeanACC -(1*SE_acc)), position = dodge, width=.25) + facet_grid(.~phase) + 
  ylab("mean ACC") + 
  coord_cartesian(ylim = c(80,100)) +
  scale_fill_grey()

```

#### 3-way ANOVA: cued trials: ACCURACY ~ 2 phase x 2 swProb x 2 trialType_sw
```{r}
z = ezANOVA(data = gpCondM1
            , dv = meanACC
            , wid = sbjId
            , within = .(phase, swProb, trialType_sw)
            , within_full = .(phase, swProb, trialType_sw)
            , type = 3
            , detailed = TRUE)
print(z)

```

#### Figure 3: Voluntary task-switching rate (VSR) ~ 2 swProb (25%, 75%)
```{r Figure3, fig.width=4.5, fig.height=4}
df <- gpCondM3 %>% select(sbjId, swProb, meanVSR)
colnames(df)[colnames(df)=="meanVSR"] <- "M"
LnMSE_VSR <- getWSSE(df)


VSR <- gpCondM3 %>%
  group_by(swProb) %>%
  summarise(gpmeanVSR = mean(meanVSR), SE_vsrRate = LnMSE_VSR)

VSR_25 = VSR[1,"gpmeanVSR"]*100
VSR_75 = VSR[2,"gpmeanVSR"]*100

print(paste("VSR_25% swProb =", sprintf("%.2f", VSR_25), "%"))
print(paste("VSR_75% swProb =", sprintf("%.2f", VSR_75), "%"))

t.test(formula = meanVSR~swProb, data = gpCondM3)

VSR  %>%
  ggplot(aes(x=swProb, y = gpmeanVSR*100)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanVSR*100 + (1*SE_vsrRate*100), ymin =gpmeanVSR*100 -(1*SE_vsrRate*100)), 
                position = dodge, width=.25) + 
  ylab("mean Switch Rate(%)") + 
  coord_cartesian(ylim = c(30,40)) +
  scale_fill_grey()


```



#### Figure 4: Voluntary task-switching rate (VSR) ~ 2 runHalf x 2 swProb (25%, 75%)
```{r Figure4, fig.width=4.5, fig.height=4}

df <- gpCondM4 %>% select(sbjId, runId_half, swProb, meanVSR)
colnames(df)[colnames(df)=="meanVSR"] <- "M"
LnMSE_runXswRate <- getWSSE(df)


RunXSwRate <- gpCondM4 %>%
  group_by(runId_half, swProb) %>% 
  summarise(gpmeanSwRate = mean(meanVSR), SE_rate = LnMSE_runXswRate) 

RunXSwRate  %>%
  ggplot(aes(x=runId_half, y = gpmeanSwRate*100, fill = swProb)) + 
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = gpmeanSwRate*100 + (1*SE_rate*100), ymin =gpmeanSwRate*100 -(1*SE_rate*100)), 
                position = dodge, width=.25) + 
  ylab("mean Switch Rate(%)") + 
  coord_cartesian(ylim = c(30,40)) +
  scale_fill_grey()


```


#### ANOVA: switch rate ~ 2 run x 2 swProb
```{r}
z = ezANOVA(data = gpCondM4
            , dv = meanVSR
            , wid = sbjId
            , within = .(runId_half, swProb)
            , within_full = .(runId_half, swProb)
            , type = 3
            , detailed = TRUE)
print(z)


paired_t_pval <-gpCondM4 %>% filter(runId_half=="SecondHalf") %>% summarise(pval_ttest = t.test(meanVSR~swProb, data=., paired = TRUE)$p.value)

print(paired_t_pval)


```